### 2025-08-05 final-presentation/final.tex
pres

\section{Proof Assistants} 
\subsection*{What are proof assistant}
\begin{frame}{What are proof assistant}
    proof assistant, are software more specifically a type of programming language thats allows us to 
    formalize mathematical proofs in computer for digital verification. 

@@

\subsection*{Why digital verification is needed?}
\begin{frame}{Need of digital verification}
    \begin{itemize}
        \item faster computation for complex problems
        \item many exceptional cases can be explored which would take mathematicians  long time \\
           \indent  \footnotesize ex:   The Kepler Conjecture's proof , which was so complex that verifying it manually would take 20 person-years, but proof assistants made this verification feasible and fast. 
    \end{itemize}
\section{Logical foundation}
\begin{frame}
    \begin{itemize}
    \item Based on logic (natural deduction, intuitionistic logic), $\lambda$-calculus, and type theory.
    \item Curry–Howard Correspondence:
    \begin{itemize}
        \item Propositions $\leftrightarrow$ Types
        \item Proofs $\leftrightarrow$ Programs
    \end{itemize}
    \item Dependently Typed Languages: Types can depend on values, enabling encoding of properties and proofs.
\end{itemize}
\subsection*{Naturl deduction}
\begin{frame}{Natural Deduction}
\begin{itemize}
    \item \textbf{Natural Deduction} is a rule-based system for deriving conclusions from assumptions in logic.
    \item Instead of using exhaustive truth tables, proofs are built step-by-step using inference rules.
    \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
    \item Basis for how proof assistants check the logical structure of proofs.
\end{itemize}
\subsection*{Ins}
\begin{frame}{Intuitionistic Logic}
\begin{itemize}
    \item \textbf{Intuitionistic Logic} formalizes constructive mathematics, where a statement is only true if a proof can be constructed.
    \item Omits some classical logic principles, such as the Law of Excluded Middle.
    \item Stronger requirement: to prove existence, a method or algorithm must be given.
    \item Proof assistants leverage this constructive approach for digital verification.
\end{itemize}
\subsection*{$\lambda$-Calculus}
\begin{frame}{$\lambda$-Calculus and Type Theory}
\begin{itemize}
    \item \textbf{$\lambda$-Calculus}: A foundational system for defining and applying functions using abstraction and application.
    \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
    \item \textbf{Curry--Howard Correspondence}: 
    \begin{itemize}
        \item Propositions $\leftrightarrow$ Types
        \item Proofs $\leftrightarrow$ Programs
    \end{itemize}
    \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.
\end{itemize}
\section{Architecture of proof assistant}
\begin{frame}{Architecture of a Proof Assistant}
\begin{itemize}
    \item \textbf{Kernel}: Minimal, trustworthy codebase enforcing logical rules and validating proofs.
    \item \textbf{Tactic Engine}: Helps build and automate proofs step by step.
    \item \textbf{Formal Proof Language}: Rigorously expresses definitions, statements, and proofs.
    \item \textbf{Libraries}: Collections of verified mathematical foundations for reuse.
    \item \textbf{User Interface}: IDEs and plugins for interactive, efficient proof development.
\end{itemize}
\subsection*{Kernel}
\begin{frame}{Kernel: The Trusted Core}
\begin{itemize}
    \item The \textbf{kernel} is the minimal and most critical part of a proof assistant.
    \item It enforces the logical rules of the underlying formal system (e.g., type theory).
    \item Responsible for \textbf{validating every proof step} to guarantee correctness.
    \item Ensures \textbf{soundness and trustworthiness}; the rest of the system depends on its integrity.
    \item Typically very small and rigorously tested or formally verified to avoid bugs.
    \item Example: Agda’s kernel is written in Haskell and integrates normalization to check definitional equality.
\end{itemize}
\subsection*{Tactic Engine}

>>>>

\section{Proof Assistants} 
\subsection*{What are proof assistant}
\begin{frame}{What are proof assistant}
    Proof assistant, are software more specifically a type of programming language thats allows us to 
    formalize mathematical proofs in computer for digital verification. 
\end{frame}
\subsection*{Why digital verification is needed?}
\begin{frame}{Need of digital verification}
    \begin{itemize}
        \item Fast and Efficient 
        \item Many cases can be explored which would take mathematicians  long time \\
           \indent  \footnotesize ex:   The Kepler Conjecture's proof , which was so complex that verifying it manually would take 20 person-years, but proof assistants made this verification feasible and fast.

        \item What if you don't use proof assistants? ABC conjecture
        \end{itemize}

@@

\begin{frame}
    \begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{pa.jpeg}
\end{figure}
\end{frame}

\section{Foundations}

\subsection*{Naturl deduction}
\begin{frame}{Natural Deduction}
\begin{itemize}
    \item \textbf{Natural Deduction} is a rule-based system for deriving conclusions from assumptions in logic.
    \item Instead of using exhaustive truth tables, proofs are built step-by-step using inference rules.
    \item Example: Proving from $A \land (A \rightarrow \bot)$ that $\bot$ (contradiction) can be derived.
    \item Basis for how proof assistants check the logical structure of proofs.
\end{itemize}
\end{frame}
\begin{frame}
    \begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{nd.jpeg}
\end{figure}
\end{frame}

\subsection*{Ins}
\begin{frame}{Intuitionistic Logic}
\begin{itemize}
    \item \textbf{Intuitionistic Logic} Also called Constructive Logic, reflects principles of constructive mathematics, where a statement is only true if a proof can be constructed.
    \item Omits some classical logic rules, such as the Law of Excluded Middle.
    \item Stronger requirement: to prove existence, a method or algorithm must be given.
    \item Proof assistants leverage this constructive approach for digital verification.
\end{itemize}
\end{frame}

\begin{frame}
    \begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{nid.jpeg}

\end{figure}
Inference Rules for Intuitionistic Logic
\end{frame}
\subsection*{$\lambda$-Calculus}
\begin{frame}{$\lambda$-Calculus and Type Theory}
\begin{itemize}
    \item \textbf{$\lambda$-Calculus}: A foundational system for defining and applying functions using abstraction and application.
    \item \textbf{Type Theory}: Assigns types to every term; ensures correctness of operations.
         \item \emph{Dependent types} allow types to depend on values, expressing complex logical properties.

    \item \textbf{Curry--Howard Correspondence}: 
    \begin{itemize}
        \item Propositions $\leftrightarrow$ Types
        \item Proofs $\leftrightarrow$ Programs
    \end{itemize}
\end{itemize}
\end{frame}
\section{Architecture of proof assistant}
\begin{frame}{Architecture of a Proof Assistant}
\begin{itemize}
    \item \textbf{Kernel}: Minimal, trustworthy codebase enforcing logical rules and validating proofs.
    \item \textbf{Tactic Engine}: Helps build and automate proofs step by step.
    \item \textbf{Formal Proof Language}: Rigorously expresses definitions, statements, and proofs.
    \item \textbf{Libraries}: Collections of verified mathematical foundations for reuse.
    \item \textbf{User Interface}: IDEs and plugins for interactive, efficient proof development.
\end{itemize}
\end{frame}
\subsection*{Kernel}
\begin{frame}{Kernel: The Trusted Core}
\begin{itemize}
    \item The \textbf{kernel} is the minimal and most critical part of a proof assistant.
    \item It enforces the logical rules of the underlying formal system (e.g., type theory).
    \item Responsible for \textbf{validating every proof step} to guarantee correctness.
    \item Ensures \textbf{soundness and trustworthiness}; the rest of the system depends on its integrity.
    \item Typically very small and rigorously tested or formally verified to avoid bugs.
    \item Example: Agda’s kernel is written in Haskell and integrates normalization to check definitional equality.
\end{itemize}
\end{frame}
\subsection*{Tactic Engine}
